Granular measures of agricultural land-use influence lake nitrogen and phosphorus differently at macroscales

### 2020-04-21 manuscript/manuscript.Rmd
add missing paragraph break
fix misreferenced table and supplement fig prefixs

Our first source for granular agricultural information was the 2010 Cropland Data Layer (CDL, \citealt{usda-nassNationalAgriculturalStatistics2019}), which we used to generate lake watershed and stream buffer estimates of the proportion of land cover due to specific crops such as corn and soybeans. We used the R package `cdlTools` to access CDL products \citep{chenToolsDownloadWork2018}. We used the 2010 CDL data because this was the first year of high resolution (30 x 30m) CDL data collection that matched our lake nutrient data collection window. Because the CDL contains detailed coverage for dozens of crops, including rare crops with little to no coverage, we re-categorized CDL data based on \citet{larkCroplandExpansionOutpaces2015} to a more limited set of categories (Table A1).

@@

Lake characteristics (e.g., maximum depth and watershed to lake area ratio) along with measures of nutrient transport (e.g., baseflow) and near-stream agriculture were significant predictors in all lake N and P models that we fit. When we  compared among different models for each nutrient individually, we found that those with agricultural watershed land-use (in the aggregate sense) were top-ranked (i.e., had the lowest absolute value leave-one-out cross validation score) for models of both TN and TP concentrations (Figure \ref{fig2, Table \ref{table2}). Although we observed no difference in the specific watershed land-use predictor used in each top-ranked model, we found differences in the extent to which each top-ranked model was substantially different from lower ranked models. For example, in the case of P, all models had nearly identical $R^2$ (0.63) and the difference between the top-ranked model and second ranked model was modest ($\Delta$ELPD = 0.41). For N models, however, agricultural and corn land-use models had higher $R^2$ (0.58) compared to other models and the difference between the top-ranked and second ranked model was more substantial ($\Delta$ELPD = 2.58, Figure \ref{fig2}, Table \ref{table1}).

@@

Both TN and TP concentrations were sensitive to measures of watershed land-use as well as near-stream agriculture (Figure \ref{fig2}, Table \ref{table1}). Despite these similarities, we found differences in both the magnitude of these effects and the extent to which we observed regional variation in watershed land-use sensitivity for N and P models (Figure \ref{fig4}). For P, there was little evidence that sensitivity to watershed land-use was regionally variable. More specifically, the credible intervals for the slope of each individual region overlapped the global slope estimate (Figure \ref{fig4}). In contrast, for N we found evidence for regional variation in sensitivity to watershed land-use. For this nutrient, the credible intervals for 2 of the 37 regions did not overlap the global estimate (Figure \ref{fig4}). These regions were found in parts of Iowa, Minnesota, and Illinois and appear to be more sensitive to watershed land-use – i.e., lake N increases at a faster rate per unit increase in agricultural land-use compared to other regions (Figure \ref{fig5}). The median soil clay content of watersheds in these regions was higher than the median across watersheds in all other regions (Figure \ref{fig6}). Furthermore, these two regions had a unique combination of both high soil clay content and extensive tile drainage (Figure S6).

@@

In our study, we found mixed evidence for regional variation in relationships with watershed land-use, depending on the lake nutrient response variable. For N, we found evidence for regional variation whereby lakes in two of the 37 regions were more sensitive to changes in agricultural land-use relative to other regions. The reasons for this elevated sensitivity are unclear, but one possible reason may be that watersheds in these more sensitive regions had higher median soil clay content than the median soil clay content of watersheds in less sensitive regions (Figure \ref{fig6}). Higher soil clay content in particular may ultimately control the nitrogen content of field runoff because it is associated with more direct (i.e., tile) drainage. For example, maps produced by \citet{capelConceptualFrameworkEffectively2018} suggest that our more sensitive regions correspond roughly with areas where field exports of nutrients are likely to bypass trapping by riparian buffers. As evidence of this association, data from \citet{nakagaki2016estimates} indicate that watersheds in these two sensitive regions had a unique combination of high soil clay content and extensive tile drainage (Figure S6).
For P, we found no evidence for regional variation in its relationship with watershed land-use. This finding, that watershed land-use can be modelled as a global (fixed) effect, is consistent with that of \citet{taranuQuantifyingRelationshipsPhosphorus2008}, who found no statistically significant differences in region-specific relationships between lake P and agricultural land-use. However, it is inconsistent with that of \citet{wagnerLandscapeDriversRegional2011}, who found regionally variable relationships between lake P and agricultural land-use using a multilevel modelling framework. One reason that both our analysis and the \citet{taranuQuantifyingRelationshipsPhosphorus2008} study did not observe regional variation in the watershed land-use versus P relationship may be that P is so strongly controlled by lake depth that there is little additional explanatory power offered by including a watershed land-use term. In addition, if lake P is controlled primarily by lake depth, which we can assume does not change with time, then our results may explain the finding of \citet{oliverUnexpectedStasisChanging2017} that lake P trends are spatially consistent whereas lake N trends have distinct regional variability.

>>>>

Our first source for granular agricultural information was the 2010 Cropland Data Layer (CDL, \citealt{usda-nassNationalAgriculturalStatistics2019}), which we used to generate lake watershed and stream buffer estimates of the proportion of land cover due to specific crops such as corn and soybeans. We used the R package `cdlTools` to access CDL products \citep{chenToolsDownloadWork2018}. We used the 2010 CDL data because this was the first year of high resolution (30 x 30m) CDL data collection that matched our lake nutrient data collection window. Because the CDL contains detailed coverage for dozens of crops, including rare crops with little to no coverage, we re-categorized CDL data based on \citet{larkCroplandExpansionOutpaces2015} to a more limited set of categories (Table S1).

@@

Lake characteristics (e.g., maximum depth and watershed to lake area ratio) along with measures of nutrient transport (e.g., baseflow) and near-stream agriculture were significant predictors in all lake N and P models that we fit. When we  compared among different models for each nutrient individually, we found that those with agricultural watershed land-use (in the aggregate sense) were top-ranked (i.e., had the lowest absolute value leave-one-out cross validation score) for models of both TN and TP concentrations (Figure \ref{fig2, Table \ref{table2}). Although we observed no difference in the specific watershed land-use predictor used in each top-ranked model, we found differences in the extent to which each top-ranked model was substantially different from lower ranked models. For example, in the case of P, all models had nearly identical $R^2$ (0.63) and the difference between the top-ranked model and second ranked model was modest ($\Delta$ELPD = 0.41). For N models, however, agricultural and corn land-use models had higher $R^2$ (0.58) compared to other models and the difference between the top-ranked and second ranked model was more substantial ($\Delta$ELPD = 2.58, Figure \ref{fig2}, Table \ref{table2}).

@@

Both TN and TP concentrations were sensitive to measures of watershed land-use as well as near-stream agriculture (Figure \ref{fig2}, Table \ref{table2}). Despite these similarities, we found differences in both the magnitude of these effects and the extent to which we observed regional variation in watershed land-use sensitivity for N and P models (Figure \ref{fig4}). For P, there was little evidence that sensitivity to watershed land-use was regionally variable. More specifically, the credible intervals for the slope of each individual region overlapped the global slope estimate (Figure \ref{fig4}). In contrast, for N we found evidence for regional variation in sensitivity to watershed land-use. For this nutrient, the credible intervals for 2 of the 37 regions did not overlap the global estimate (Figure \ref{fig4}). These regions were found in parts of Iowa, Minnesota, and Illinois and appear to be more sensitive to watershed land-use – i.e., lake N increases at a faster rate per unit increase in agricultural land-use compared to other regions (Figure \ref{fig5}). The median soil clay content of watersheds in these regions was higher than the median across watersheds in all other regions (Figure \ref{fig6}). Furthermore, these two regions had a unique combination of both high soil clay content and extensive tile drainage (Figure S6).

@@

In our study, we found mixed evidence for regional variation in relationships with watershed land-use, depending on the lake nutrient response variable. For N, we found evidence for regional variation whereby lakes in two of the 37 regions were more sensitive to changes in agricultural land-use relative to other regions. The reasons for this elevated sensitivity are unclear, but one possible reason may be that watersheds in these more sensitive regions had higher median soil clay content than the median soil clay content of watersheds in less sensitive regions (Figure \ref{fig6}). Higher soil clay content in particular may ultimately control the nitrogen content of field runoff because it is associated with more direct (i.e., tile) drainage. For example, maps produced by \citet{capelConceptualFrameworkEffectively2018} suggest that our more sensitive regions correspond roughly with areas where field exports of nutrients are likely to bypass trapping by riparian buffers. As evidence of this association, data from \citet{nakagaki2016estimates} indicate that watersheds in these two sensitive regions had a unique combination of high soil clay content and extensive tile drainage (Figure S6).

For P, we found no evidence for regional variation in its relationship with watershed land-use. This finding, that watershed land-use can be modelled as a global (fixed) effect, is consistent with that of \citet{taranuQuantifyingRelationshipsPhosphorus2008}, who found no statistically significant differences in region-specific relationships between lake P and agricultural land-use. However, it is inconsistent with that of \citet{wagnerLandscapeDriversRegional2011}, who found regionally variable relationships between lake P and agricultural land-use using a multilevel modelling framework. One reason that both our analysis and the \citet{taranuQuantifyingRelationshipsPhosphorus2008} study did not observe regional variation in the watershed land-use versus P relationship may be that P is so strongly controlled by lake depth that there is little additional explanatory power offered by including a watershed land-use term. In addition, if lake P is controlled primarily by lake depth, which we can assume does not change with time, then our results may explain the finding of \citet{oliverUnexpectedStasisChanging2017} that lake P trends are spatially consistent whereas lake N trends have distinct regional variability.
